Ecuador plant trait TRY Cleaning

### 2023-07-13 archive/L0_NorthAndes_plants_GBIF.Rmd
move old code to archive

# Install and load required packages
```{r}
library(rgbif)
library(DataExplorer)
library(summarytools)
library(bdvis)
library(visdat)
library(dplyr)
```

# Set GBIF User information
```{r}
#only run once
#install.packages("usethis")
#usethis::edit_r_environ()
```
# Get Taxon Key for Plants
```{r}
name_backbone("Plantae")
```

```{r}
occ_download(
pred("taxonKey", 6), 
pred("hasGeospatialIssue", FALSE),
pred("hasCoordinate", TRUE),
pred_in("country",c("EC","CO","VE")), 
pred_not(pred_in("basisOfRecord",c("FOSSIL_SPECIMEN","LIVING_SPECIMEN"))),
format = "SIMPLE_CSV"
)
```
# Retrive GBIF download, save file, and load into r
```{r}
d <- occ_download_get("0282867-220831081235567", path = output_path) %>%
  occ_download_import()
```

@@

```{r}
# using summarytools package
# ERROR -- Error in hist.default(data, breaks = breaks_x, plot = FALSE) :
# some 'x' not counted; maybe 'breaks' do not span range of 'x'
#d_summary <- dfSummary(d, 
#            na.col       = FALSE,
#            style        = "multiline",
#            plain.ascii  = TRUE,
#            graph.magnif = .8)
#print(d_summary, method = "render")
#view(d_summary, file = file.path(output_path,"Ecuador_BIEN_traits_summary.html"))
```

@@

```{r}
## ERROR - Error in !dbPreExists : invalid argument type
## ERROR - Error in postgresqlNewConnection(drv, ...) :
# RPosgreSQL error: could not connect postgres@localhost:5432 on dbname "test": could not connect to server: Connection refused (0x0000274D/10061)
#bdcalendarheat(d_vis)
#distrigraph(d_vis, ptype = "species")
#distrigraph(d_vis, ptype = "efforts")
#distrigraph(d_vis, ptype = "cell")
#distrigraph(d_vis, ptype = "effortspecies")
#bdcomplete(d_vis)
```

@@

# Write data to csv
```{r}
NorthAndes_GBIF_path <- file.path(output_path,"NorthAndes_GBIF_occurences.csv") 
write.csv(d, file = NorthAndes_GBIF_path)
```

>>>>

# Install and load required packages
```{r}
library(rgbif)
library(DataExplorer)
library(bdvis)
library(visdat)
library(dplyr)
```

# Set GBIF User information
```{r}
#only run once
#install.packages("usethis")
#usethis::edit_r_environ()
```

# Get Taxon Key for Plants
```{r}
name_backbone("Plantae")
```

```{r}
occ_download(
pred("taxonKey", 6), 
pred("hasGeospatialIssue", FALSE),
pred("hasCoordinate", TRUE),
pred_in("country",c("EC","CO","VE")), 
pred_not(pred_in("basisOfRecord",c("FOSSIL_SPECIMEN","LIVING_SPECIMEN"))),
format = "SIMPLE_CSV"
)
```

# Retrive GBIF download, save file, and load into r
```{r}
d <- occ_download_get("0282867-220831081235567", path = output_path) %>%
  occ_download_import()
```

@@

@@

@@

# Write data to csv
```{r}
NorthAndes_GBIF_path <- file.path(output_path,"NorthAndes_GBIF_plant_occ.csv") 
write.csv(d, file = NorthAndes_GBIF_path)
```

### 2023-07-13 archive/L0_NorthAndes_plants_IUCN.Rmd
move old code to archive

# Read in IUCN data
Data was obtained from https://www.iucnredlist.org/search?dl=true&permalink=a44f4942-2b56-47b9-8122-9896ac4af647
```{r}
assessments <- read.csv(file.path(data_path,"IUCN/Northern Andes 2 - Search Results/assessments.csv"))
common_names <- read.csv(file.path(data_path,"IUCN/Northern Andes 2 - Search Results/common_names.csv"))
conservation_needed <- read.csv(file.path(data_path,"IUCN/Northern Andes 2 - Search Results/conservation_needed.csv"))
countries <- read.csv(file.path(data_path,"IUCN/Northern Andes 2 - Search Results/countries.csv"))
credits <- read.csv(file.path(data_path,"IUCN/Northern Andes 2 - Search Results/credits.csv"))
dois <- read.csv(file.path(data_path,"IUCN/Northern Andes 2 - Search Results/dois.csv"))
fao <- read.csv(file.path(data_path,"IUCN/Northern Andes 2 - Search Results/fao.csv"))
habitats <- read.csv(file.path(data_path,"IUCN/Northern Andes 2 - Search Results/habitats.csv"))
plant_specific <- read.csv(file.path(data_path,"IUCN/Northern Andes 2 - Search Results/plant_specific.csv"))
references <- read.csv(file.path(data_path,"IUCN/Northern Andes 2 - Search Results/references.csv"))
research_needed <- read.csv(file.path(data_path,"IUCN/Northern Andes 2 - Search Results/research_needed.csv"))
simple_summary <- read.csv(file.path(data_path,"IUCN/Northern Andes 2 - Search Results/simple_summary.csv"))
synonyms <- read.csv(file.path(data_path,"IUCN/Northern Andes 2 - Search Results/synonyms.csv"))
taxonomy <- read.csv(file.path(data_path,"IUCN/Northern Andes 2 - Search Results/taxonomy.csv"))
threats <- read.csv(file.path(data_path,"IUCN/Northern Andes 2 - Search Results/threats.csv"))
usetrade <- read.csv(file.path(data_path,"IUCN/Northern Andes 2 - Search Results/usetrade.csv"))
points_data <- read.csv(file.path(data_path,"IUCN/Northern Andes 3 - Range data Points/points_data.csv"))
```

>>>>

# Read in IUCN data
Data was obtained from https://www.iucnredlist.org/search?dl=true&permalink=a44f4942-2b56-47b9-8122-9896ac4af647
```{r}
assessments <- read.csv(file.path(data_path,"IUCN/Northern Andes 1 - Search Results/assessments.csv"))
common_names <- read.csv(file.path(data_path,"IUCN/Northern Andes 1 - Search Results/common_names.csv"))
conservation_needed <- read.csv(file.path(data_path,"IUCN/Northern Andes 1 - Search Results/conservation_needed.csv"))
countries <- read.csv(file.path(data_path,"IUCN/Northern Andes 1 - Search Results/countries.csv"))
credits <- read.csv(file.path(data_path,"IUCN/Northern Andes 1 - Search Results/credits.csv"))
dois <- read.csv(file.path(data_path,"IUCN/Northern Andes 1 - Search Results/dois.csv"))
fao <- read.csv(file.path(data_path,"IUCN/Northern Andes 1 - Search Results/fao.csv"))
habitats <- read.csv(file.path(data_path,"IUCN/Northern Andes 1 - Search Results/habitats.csv"))
plant_specific <- read.csv(file.path(data_path,"IUCN/Northern Andes 1 - Search Results/plant_specific.csv"))
references <- read.csv(file.path(data_path,"IUCN/Northern Andes 1 - Search Results/references.csv"))
research_needed <- read.csv(file.path(data_path,"IUCN/Northern Andes 1 - Search Results/research_needed.csv"))
simple_summary <- read.csv(file.path(data_path,"IUCN/Northern Andes 1 - Search Results/simple_summary.csv"))
synonyms <- read.csv(file.path(data_path,"IUCN/Northern Andes 1 - Search Results/synonyms.csv"))
taxonomy <- read.csv(file.path(data_path,"IUCN/Northern Andes 1 - Search Results/taxonomy.csv"))
threats <- read.csv(file.path(data_path,"IUCN/Northern Andes 1 - Search Results/threats.csv"))
usetrade <- read.csv(file.path(data_path,"IUCN/Northern Andes 1 - Search Results/usetrade.csv"))
points_data <- read.csv(file.path(data_path,"IUCN/Northern Andes 3 - Range data Points/points_data.csv"))
```

### 2023-07-13 archive/L0_NorthAndes_plants_TRY.Rmd
move old code to archive

# Read in TRY data
Data was obtained from https://www.try-db.org/TryWeb/Prop0.php 
```{r}
data <- rtry_import("G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0/TRY/TRY_Ecuador_PublicData_10192020.txt", encoding = "Latin-1")
data <- as.data.frame(data)
```

>>>>

# Read in TRY data
Data was obtained from https://www.try-db.org/TryWeb/Prop0.php 
```{r}
data <- rtry_import("G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0/TRY/23650_24112022083128/23650.txt", encoding = "Latin-1")
data <- as.data.frame(data)
```

### 2023-07-13 archive/L1_NorthAndes_plants_TRY.Rmd
move old code to archive

# Configure TRY data to Darwin Core Standards
```{r}
# darwinizing a dataset
config <- read.csv(file.path(output_path,"Ecuador_plant_TRY_trait_ConfigurationTable"))

@@

# Pre-filter
```{r}
# Flag records missing species names
check_pf <-
  bdc_scientificName_empty(
  data = database,
  sci_name = "scientificName")
```

@@

# Flagging common time issues
```{r}
check_time <-
  bdc_eventDate_empty(data = database, eventDate = "verbatimEventDate")
```

@@

# Flagging common spatial issues
```{r}
check_space <-
  bdc_coordinates_precision(
    data = database,
    lon = "decimalLongitude",
    lat = "decimalLatitude",
    ndec = c(0, 1) # number of decimals to be tested
  )
```

@@

# Clean and parse species names
```{r}
parse_names <-
  bdc_clean_names(sci_names = database$scientificName, save_outputs = FALSE)
```

@@

# Transform to wide table
```{r}
#-------------------------------------------------
# Exclude
# 1. All entries with "" in TraitID
# 2. Potential categorical traits that don't have a StdValue
# 3. Traits that have not yet been standardized in TRY
# Then select the relevant columns for transformation
# Note: The complete.cases() is used to ensure the cases are complete,
#       i.e. have no missing values
#-------------------------------------------------
num_traits <- rtry_select_row(workdata, complete.cases(TraitID) & complete.cases(StdValue))
num_traits <- rtry_select_col(num_traits, ObservationID, AccSpeciesID, AccSpeciesName, TraitID, TraitName, StdValue, UnitName)
# Extract the unique value of latitude (DataID 59) and the corresponding ObservationID
workdata_lat <- rtry_select_anc(try_data, 59)

>>>>

# Configure TRY data to Darwin Core Standards
```{r}
# darwinizing a dataset
config <- read.csv(file.path(output_path,"Ecuador_plant_TRY_trait_ConfigurationTable.csv"))

@@

# Pre-filter
Modified from https://brunobrr.github.io/bdc/articles/prefilter.html
```{r}
# Flag records missing species names
check_pf <-
  bdc_scientificName_empty(
  data = database,
  sci_name = "scientificName")
```

@@

# Flagging common time issues
Modified from https://brunobrr.github.io/bdc/articles/time.html
```{r}
check_time <-
  bdc_eventDate_empty(data = database, eventDate = "verbatimEventDate")
```

@@

# Flagging common spatial issues
Modified from https://brunobrr.github.io/bdc/articles/space.html
```{r}
check_space <-
  bdc_coordinates_precision(
    data = database,
    lon = "decimalLongitude",
    lat = "decimalLatitude",
    ndec = c(0, 1) # number of decimals to be tested
  )
```

@@

# Clean and parse species names
Modified from https://brunobrr.github.io/bdc/articles/taxonomy.html
```{r}
parse_names <-
  bdc_clean_names(sci_names = database$scientificName, save_outputs = FALSE)
```

@@

# Transform to wide table
Modified from https://cran.r-project.org/web/packages/rtry/vignettes/rtry-workflow-general.html
```{r}
#-------------------------------------------------
# Exclude
# 1. All entries with "" in TraitID
# 2. Potential categorical traits that don't have a StdValue
# 3. Traits that have not yet been standardized in TRY
# Then select the relevant columns for transformation
# Note: The complete.cases() is used to ensure the cases are complete,
#       i.e. have no missing values
#-------------------------------------------------
num_traits <- rtry_select_row(workdata, complete.cases(TraitID) & complete.cases(StdValue))
num_traits <- rtry_select_col(num_traits, ObservationID, AccSpeciesID, AccSpeciesName, TraitID, TraitName, StdValue, UnitName)
# Extract the unique value of latitude (DataID 59) and the corresponding ObservationID
workdata_lat <- rtry_select_anc(try_data, 59)
